Identifying Inter-Genomic Repeats Using Fast, Approximate Methods for Betweenness Centrality

### 2015-05-18 manuscript/manuscript.tex
edited IQR

Decreasing the betweenness centrality cutoff requirement results in a trade-off between the different statistical measures.
Instead of using the default cutoff of $3\sigma$ suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
In the 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the IQR cutoff method.
Conversely, the specificity and and precision drops from 99.3\% to 90.9\% and 25\% to 12.5\%, respectively.

@@

The advantage of using the node sampling algorithm is that it appears to scale linearly in practice.
Thus, if runtime is a concern, small increasing subsets of nodes can be sampled and the runtime of a larger sample can be extrapolated.
We can pair this information with how the subset of nodes marked as inter-genomic repeats changes as we increase the sample size.
If the same set of nodes are being marked as inter-genomic repeats, then the algorithm can be terminated before sampling a larger set of nodes.
Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled until the results stabilize.
This is the main advantage of the shortest path sampling approach.
While the shortest path sampling approach takes a fair bit longer in our evaluation then sampling a small number of nodes, it has a probabilistic guarantee associated with the betweenness centrality computation.

>>>>

Decreasing the betweenness centrality cutoff requirement results in a trade-off between the different statistical measures.
We observed that for 10-genome assembly graph, the sensitivity was very less according to the cut off criteria in Bambus2. This was due to the fact that cut off was strict and was eliminating actual repeats.
Instead of using the cut off criteria suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
When we use this criteria on 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the IQR cutoff method.
Conversely, the specificity and and precision drops from 99.3\% to 90.9\% and 25\% to 12.5\%, respectively.
We could achieve better specificity at by not losing much on sensitivity.

@@

The advantage of using the node sampling algorithm is that it appears to scale linearly in practice.
Thus, if runtime is a concern, small increasing subsets of nodes can be sampled and the runtime of a larger sample can be extrapolated.
We can pair this information with how the subset of nodes marked as inter-genomic repeats changes as we increase the sample size.
If the same set of nodes are being marked as inter-genomic repeats, then the algorithm can be terminated before sampling a larger set of nodes.
Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled.
This is the main advantage of the shortest path sampling approach.
While the shortest path sampling approach takes a fair bit longer in our approach then small node sample sizes, it has a probabilistic guarantee associated with the betweenness centrality computation.
